07-logistic_regression

### 2019-02-22 journals/01-training_and_testing_data/01-training_and_testing_data.Rmd
created folders for all journals and re-knit all journals

This journal includes the code that is used to clean the training and testing datasets.

@@

```{r}
# Obtain features used when fitting the rtrees random forest
rf_features <- rownames(bulletr::rtrees$importance)
rf_features
```

The definitions of these features are inclued below.

@@

```{r}
# Load in the Hamby 173 and 252 dataset
hamby173and252_raw <- read.csv("../data/raw/features-hamby173and252.csv") %>%
  filter(study1 != "Cary", study2 != "Cary") %>%
  mutate(study1 = factor(study1), 
         study2 = factor(study2))
```

@@

I discovered that rhe number of rtrees predictions does not match the length of my training data. I need to ask Heike about this.

@@

```{r}
# Remove the comparisons involving the four lands with tank rash
hamby173and252_train <- hamby173and252_train_cleaning %>%
  mutate(bbl1 = barrel1:bullet1:land1,
         bbl2 = barrel2:bullet2:land2) %>%
  filter(!(bbl1 %in% c("6:2:1", "9:2:4", "Unknown:B:2", "Unknown:Q:4") | 
           bbl2 %in% c("6:2:1", "9:2:4", "Unknown:B:2", "Unknown:Q:4")))
```

@@

```{r}
# Save the datasets and response variables as .csv files
write.csv(hamby173and252_train, "../data/hamby173and252_train.csv", row.names = FALSE)
```

@@

```{r}
# Load in the Hamby 224 datasets
hamby224_set1 <- readRDS("../data/raw/h224-set1-features.rds")
hamby224_set11 <- readRDS("../data/raw/h224-set11-features.rds")
```

@@

```{r}
# Save the test data as a .csv file
write.csv(hamby224_test, "../data/hamby224_test.csv", row.names = FALSE)
```

>>>>

This journal includes a description of the Hamby study datasets, and the code that is used to clean the Hamby training and testing datasets.

@@

```{r}
# Obtain features used when fitting the rtrees random forest
rf_features <- rownames(bulletr::rtrees$importance)
```

# Description of the Hamby Data

The Hamby data is based on several test sets of bullets from the study described in the paper "The Identification of Bullets Fired from 10 Consecutively Rifled 9mm Ruger Pistol Barrels: A Research Project Involving 507 Participants from 20 Countries" by James E. Hamby Et. Al. In this study, sets of bullets from both "known" and "unknown" gun barrels were sent to firearm examiners around the world. The examiners were asked to use the known bullets to identify which barrels the unknown bullets came from.

The test sets were created using 1 pistol and 10 barrels. Each test set contains a total of 35 bullets, which are made up of 20 known bullets and 15 unknown bullets. The 20 known bullets were created by firing two bullets from each of the 10 barrels. These are referred to as known bullets, because when they were sent to the firearm examiners, the barrel number that each bullet was fired from was listed with the bullet. The 15 unknown bullets were created by firing 15 bullets in some manner from the 10 barrels such that at least one unknown bullet came from each barrel and no more than three unknown bullets came from the same barrel. These are referred to as the unknown bullets, because when they were sent to the firearm examiners, the barrel number that the bullet was fired from was not listed with the bullet. A total of 240 test sets were created for the study.

CSAFE has access to test sets 44, 173, and 252. The bullets were scanned 6 times using a high powered microscope to obtain an image of each of the 6 lands from the bullets. The scans for test sets 173 and 252 were done by NIST, and the scans for test set 44 was done by CSAFE. The data from these images were processed to obtain a signature associated with each land. The paper "Automatic Matching of Bullet Land Impressions" by Hare, Hofmann, and Carriquiry (https://arxiv.org/abs/1601.05788) provides more descriptions of how the signatures were obtained.

CSAFE aggregated the signatures from test sets 173 and 252 and left the signatures from test set 44 separate. Within these two groups, pairs of lands were evaluated to determine how similar the signatures from the two lands were. This was done by measuring a set of variables they determined that would capture how alike the two signatures where. Some of these variables are described in Hare, Hofmann, and Carriquiry. The vignette at https://github.com/heike/bulletxtrctr/blob/master/vignettes/features.Rmd includes some additional descriptions. Note that it was not possible to evaluate all pairs of lands due to tank rash on some of the lands. The data set created from the comparisons of test sets 173 and 252 will be referred to as `hamby173and252` throughout these journals. Hare, Hofmann, and Carriquiry used `hamby173and252` as a training data set to fit the random forest model `rtrees`. The data set created from the comparisons of test sets 44 will be referred to as `hamby44` and is not used in this research project.

### Variable Definitions

The definitions of the variables...

@@

```{r}
# Load in the Hamby 173 and 252 dataset
hamby173and252_raw <- read.csv("../../data/raw/features-hamby173and252.csv") %>%
  filter(study1 != "Cary", study2 != "Cary") %>%
  mutate(study1 = factor(study1), 
         study2 = factor(study2))
```

@@

I discovered that the number of rtrees predictions does not match the length of my training data. I need to ask Heike about this.

@@

```{r}
# Remove the comparisons involving the four lands with tank rash
hamby173and252_train <- hamby173and252_train_cleaning %>%
  mutate(bbl1 = barrel1:bullet1:land1,
         bbl2 = barrel2:bullet2:land2) %>%
  filter(!(bbl1 %in% c("6:2:1", "9:2:4", "Unknown:B:2", "Unknown:Q:4") | 
           bbl2 %in% c("6:2:1", "9:2:4", "Unknown:B:2", "Unknown:Q:4"))) %>%
  select(-bbl1, -bbl2)
```

@@

```{r}
# Save the datasets and response variables as .csv files
write.csv(hamby173and252_train, "../../data/hamby173and252_train.csv", row.names = FALSE)
```

@@

```{r}
# Load in the Hamby 224 datasets
hamby224_set1 <- readRDS("../../data/raw/h224-set1-features.rds")
hamby224_set11 <- readRDS("../../data/raw/h224-set11-features.rds")
```

@@

```{r}
# Save the test data as a .csv file
write.csv(hamby224_test, "../../data/hamby224_test.csv", row.names = FALSE)
```

### 2019-02-22 journals/02-plotting_training_and_testing_data/02-plotting_training_and_testing_data.Rmd
created folders for all journals and re-knit all journals

---
title: "02-plotting_features"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

@@

```{r}
# Load in the training data (Hamby Data 173 and 252)
hamby173and252_train <- read.csv("../data/hamby173and252_train.csv")

# Load in the testing data (Hamby Data 224 Sets 1 and 11)
hamby224_test <- read.csv("../data/hamby224_test.csv")
```

>>>>

---
title: "02-plotting_training_and_testing_data"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

@@

```{r}
# Load in the training data (Hamby Data 173 and 252)
hamby173and252_train <- read.csv("../../data/hamby173and252_train.csv")

# Load in the testing data (Hamby Data 224 Sets 1 and 11)
hamby224_test <- read.csv("../../data/hamby224_test.csv")
```

### 2019-02-22 journals/03-computation_issues/03-computation_issues.Rmd
created folders for all journals and re-knit all journals

```{r}
# Load libraries
library(tidyverse)
library(plotly)
library(caret)
library(randomForest)
library(lime)
```

@@

library(tictoc)
library(future)
library(furrr)

>>>>

```{r}
# Load libraries
library(caret)
library(lime)
library(plotly)
library(randomForest)
library(tidyverse)
```

@@

library(furrr)
library(future)
library(tictoc)

### 2019-02-22 journals/04-lime_algorithm_and_proposed_methods/04-lime_algorithm_and_proposed_methods.Rmd
created folders for all journals and re-knit all journals

# Source functions
source("../code/helper_functions.R")
```

@@

```{r}
# Load in the training data (Hamby Data 173 and 252)
hamby173and252_train <- read.csv("../data/hamby173and252_train.csv")
```

@@

@@

# Save the subsampled data
write.csv(hamby173and252_train_sub, "../data/hamby173and252_train_sub.csv", row.names = FALSE)

>>>>

# Source functions
source("../../code/helper_functions.R")
```

@@

```{r}
# Load in the training data (Hamby Data 173 and 252)
hamby173and252_train <- read.csv("../../data/hamby173and252_train.csv")
```

@@

# Inputs for LIME Functions

The default input settings for the `lime` and `explain` functions in the lime package are described below.

### `lime` function

The default settings for the `lime` function are as follows.

- `bin_continuous = TRUE`: If set to `TRUE`, the continuous variables will be binned when making the explanations. If they are not binned, then perturbations will be obtained by either simulating using kernel density estimation or a normal distribution depending on what the option of `use_density` is set to. 
- `bins = 4`: The number of bins to divide the continuous variables into. The default is 4 bins.
- `quantile_bins = TRUE`: If set to `TRUE`, the bins will be be based on `n_bins` quantiles. Otherwise, the bins will be spread evenly over the range of the training data.
- `use_density = TRUE`: If `bin_continuous` is set to `FALSE` and this is set to `TRUE`, then the continuous data will be sampled using kernel density estimation. Otherwise, it will be assumed that the continuous features follow a normal distribution and samples will be drawn from a normal distribution with the mean and standard deviation set to the sample mean and standard deviation associated with the feature.

### `explain` function

For our research project, the number of labels (`n_labels`) for the response variable will be set to 1 and the number of features to include in the explanations (`n_features`) will be set to 2. The default values for the other options are as follows.

- `n_permutations = 5000`: The number of perturbations generated for each feature.
- `feature_select = 'auto'`: This is the feature selection method for choosing the number of features specified. This uses forward selection if $m\le 6$ and otherwise highest weights. The other feature selection options are as follows.
    - `none`: Use all features for the explanation. Not advised unless you have very few features.
    - `forward selection`: Features are added one by one based on their improvements to a ridge regression fit of the complex model outcome.
    - `highest weights`: The m features with highest absolute weight in a ridge regression fit of the complex model outcome are chosen.
    - `lasso`: The m features that are least prone to shrinkage based on the regularization path of a lasso fit of the complex model outcome is chosen.
    - `tree`: A tree is fitted with log2(m) splits, to use at max m features. It may possibly select less.
- `dist_fun = 'gower'`: The distance function to be used for determining how close a perturbation is to the test point.
- `kernel_width = NULL`: If `dist_fun` is not set to `gower`, then this is the kernel width that will be used in the distance function.

@@

# Save the subsampled data
write.csv(hamby173and252_train_sub, "../../data/hamby173and252_train_sub.csv", row.names = FALSE)

### 2019-02-22 journals/05-applying_lime_to_rtrees/05-applying_lime_to_rtrees.Rmd
created folders for all journals and re-knit all journals

---
title: "05-applying_lime"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

@@

This journal includes the code for applying lime to the random forest fit to the bullet data.

@@

# Load rtrees into session
data("rtrees")

# Source functions
source("../code/helper_functions.R")
```

@@

```{r}
# Load in the training data (Hamby Data 173 and 252)
hamby173and252_train <- read.csv("../data/hamby173and252_train.csv")

# Load in the testing data (Hamby Data 224 Sets 1 and 11)
hamby224_test <- read.csv("../data/hamby224_test.csv")
```

@@

```{r}
# Apply the run_lime function if the lime results file does not already exist
if(!file.exists("../data/hamby224_lime_inputs.rds")) {

@@

  # Save the lime objects
  saveRDS(hamby224_lime_inputs, "../data/hamby224_lime_inputs.rds")
  saveRDS(hamby224_lime, "../data/hamby224_lime.rds")
  saveRDS(hamby224_explain, "../data/hamby224_explain.rds")

@@

  # Load in the lime objects
  hamby224_lime_inputs <- readRDS("../data/hamby224_lime_inputs.rds")
  hamby224_lime <- readRDS("../data/hamby224_lime.rds")
  hamby224_explain <- readRDS("../data/hamby224_explain.rds")

@@

```{r}
# Create a dataframe with the bins
if (!file.exists("../data/hamby224_bins.csv")) {

@@

  # Save the bin boundaries and bins
  saveRDS(hamby224_bin_boundaries, "../data/hamby224_bin_boundaries.rds")
  saveRDS(hamby224_bins, "../data/hamby224_bins.rds")

@@

  # Load in the bin boundaries and bins
  hamby224_bin_boundaries <- readRDS("../data/hamby224_bin_boundaries.rds")
  hamby224_bins <- readRDS("../data/hamby224_bins.rds")

@@

```{r}
# Create the test_explain combined data if the file does not already exist
if(!file.exists("../data/hamby224_test_explain.rds")) {

@@

  # Save the combined test and explain data
  saveRDS(hamby224_test_explain, "../data/hamby224_test_explain.rds")

@@

  # Load in the data
  hamby224_test_explain <- readRDS("../data/hamby224_test_explain.rds")

@@

```{r}
# Create the lime comparison data if the file does not already exist
if(!file.exists("../data/hamby224_lime_comparisons.rds")) {

@@

  # Save the lime comparison data frame
  saveRDS(hamby224_lime_comparisons, "../data/hamby224_lime_comparisons.rds")

@@

  # Load in the lime comparison data frame
  hamby224_lime_comparisons <- readRDS("../data/hamby224_lime_comparisons.rds")

@@

```{r}
# Perform the sensitivity analysis if not already saved
if(!file.exists("../data/hamby224_sensitivity_inputs.rds")) {

@@

  # Save the sensitivity inputs and outputs
  saveRDS(hamby224_sensitivity_inputs, "../data/hamby224_sensitivity_inputs.rds")
  saveRDS(hamby224_sensitivity_outputs, "../data/hamby224_sensitivity_outputs.rds")

@@

  # Load in the sensitivity inputs and outputs
  hamby224_sensitivity_inputs <- readRDS("../data/hamby224_sensitivity_inputs.rds")
  hamby224_sensitivity_outputs <- readRDS("../data/hamby224_sensitivity_outputs.rds")

@@

```{r}
# Join the sensativity outputs with the test data if not already saved
if(!file.exists("../data/hamby224_sensitivity_joined.rds")) {

@@

  # Save the joined test and explained dataframes
  saveRDS(hamby224_sensitivity_joined, "../data/hamby224_sensitivity_joined.rds")

@@

  # Load in the sensitivity inputs and outputs
  hamby224_sensitivity_joined <- readRDS("../data/hamby224_sensitivity_joined.rds")

>>>>

---
title: "05-applying_lime_to_rtrees"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

@@

This journal includes a description of the random forest model `rtrees` fit to the bullet data by Eric and the code for applying lime to `rtrees`.

@@

# Source functions
source("../../code/helper_functions.R")
```

@@

```{r}
# Load in the training data (Hamby Data 173 and 252)
hamby173and252_train <- read.csv("../../data/hamby173and252_train.csv")

# Load in the testing data (Hamby Data 224 Sets 1 and 11)
hamby224_test <- read.csv("../../data/hamby224_test.csv")
```

# Description of `rtrees`

The random forest model that was fit to the Hamby training dataset is discussed in the paper "Automatic Matching of Bullet Land Impressions" by E. Hare Et. Al. The model is available through the GitHub version of the `bulletr` package. The link to the GitHub repository for `bulletr` is https://github.com/CSAFE-ISU/bulletr.

The random forest model is stored in the package `bulletr` as a data object called `rtrees`. The paper does not include many details about the model fitting process, so I looked at some of the model features, which are printed below. These values told me that the model was fit using the predictive features of `ccf`, `rough_cor`, `D`, `sd_D`, `matches`, `mismatches`, `cms`, `non_cms`, and `sum_peaks` with 300 trees and the parameter of `mtry` set to 3, which is the "number of variables randomly sampled as candidates at each split" as explained in the `randomforest` package documentation.

```{r}
# Importance values from the random forest
rtrees$importance

# The number of trees used in the random forest
rtrees$ntree

# "Number of variables randomly sampled as candidates at each split"
rtrees$mtry
```

@@

```{r}
# Apply the run_lime function if the lime results file does not already exist
if(!file.exists("../../data/hamby224_lime_inputs.rds")) {

@@

  # Save the lime objects
  saveRDS(hamby224_lime_inputs, "../../data/hamby224_lime_inputs.rds")
  saveRDS(hamby224_lime, "../../data/hamby224_lime.rds")
  saveRDS(hamby224_explain, "../../data/hamby224_explain.rds")

@@

  # Load in the lime objects
  hamby224_lime_inputs <- readRDS("../../data/hamby224_lime_inputs.rds")
  hamby224_lime <- readRDS("../../data/hamby224_lime.rds")
  hamby224_explain <- readRDS("../../data/hamby224_explain.rds")

@@

```{r}
# Create a dataframe with the bins
if (!file.exists("../../data/hamby224_bins.csv")) {

@@

  # Save the bin boundaries and bins
  saveRDS(hamby224_bin_boundaries, "../../data/hamby224_bin_boundaries.rds")
  saveRDS(hamby224_bins, "../../data/hamby224_bins.rds")

@@

  # Load in the bin boundaries and bins
  hamby224_bin_boundaries <- readRDS("../../data/hamby224_bin_boundaries.rds")
  hamby224_bins <- readRDS("../../data/hamby224_bins.rds")

@@

```{r}
# Create the test_explain combined data if the file does not already exist
if(!file.exists("../../data/hamby224_test_explain.rds")) {

@@

  # Save the combined test and explain data
  saveRDS(hamby224_test_explain, "../../data/hamby224_test_explain.rds")

@@

  # Load in the data
  hamby224_test_explain <- readRDS("../../data/hamby224_test_explain.rds")

@@

```{r}
# Create the lime comparison data if the file does not already exist
if(!file.exists("../../data/hamby224_lime_comparisons.rds")) {

@@

  # Save the lime comparison data frame
  saveRDS(hamby224_lime_comparisons, "../../data/hamby224_lime_comparisons.rds")

@@

  # Load in the lime comparison data frame
  hamby224_lime_comparisons <- readRDS("../../data/hamby224_lime_comparisons.rds")

@@

```{r}
# Perform the sensitivity analysis if not already saved
if(!file.exists("../../data/hamby224_sensitivity_inputs.rds")) {

@@

  # Save the sensitivity inputs and outputs
  saveRDS(hamby224_sensitivity_inputs, "../../data/hamby224_sensitivity_inputs.rds")
  saveRDS(hamby224_sensitivity_outputs, "../../data/hamby224_sensitivity_outputs.rds")

@@

  # Load in the sensitivity inputs and outputs
  hamby224_sensitivity_inputs <- readRDS("../../data/hamby224_sensitivity_inputs.rds")
  hamby224_sensitivity_outputs <- readRDS("../../data/hamby224_sensitivity_outputs.rds")

@@

```{r}
# Join the sensativity outputs with the test data if not already saved
if(!file.exists("../../data/hamby224_sensitivity_joined.rds")) {

@@

  # Save the joined test and explained dataframes
  saveRDS(hamby224_sensitivity_joined, "../../data/hamby224_sensitivity_joined.rds")

@@

  # Load in the sensitivity inputs and outputs
  hamby224_sensitivity_joined <- readRDS("../../data/hamby224_sensitivity_joined.rds")

### 2019-02-22 journals/06-assessing_lime_explanations/06-assessing_lime_explanations.Rmd
created folders for all journals and re-knit all journals

# Load in the testing data (Hamby Data 224 Sets 1 and 11)
hamby224_test <- read.csv("../data/hamby224_test.csv") # created in the paper

# Read in the test_explain data
hamby224_test_explain <- readRDS("../data/hamby224_test_explain.rds")

hamby224_sensitivity_joined <- readRDS("../data/hamby224_sensitivity_joined.rds")

@@

```{r}
# Read in the lime comparison data (the comparisons are run in the firearm examiner
# paper since they will be discussed in the paper)
hamby224_lime_comparisons <- readRDS("../data/hamby224_lime_comparisons.rds")
```

>>>>

# Load in the testing data (Hamby Data 224 Sets 1 and 11)
hamby224_test <- read.csv("../../data/hamby224_test.csv") # created in the paper

# Read in the test_explain data
hamby224_test_explain <- readRDS("../../data/hamby224_test_explain.rds")

hamby224_sensitivity_joined <- readRDS("../../data/hamby224_sensitivity_joined.rds")

@@

```{r}
# Read in the lime comparison data (the comparisons are run in the firearm examiner
# paper since they will be discussed in the paper)
hamby224_lime_comparisons <- readRDS("../../data/hamby224_lime_comparisons.rds")
```

### 2019-02-22 journals/07-logistic_regression/07-logistic_regression.Rmd
created folders for all journals and re-knit all journals

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, eval = TRUE)
```

@@

```{r}
# Load libraries
library(tidyverse)
library(lime)
```

@@

```{r}
# Fit a logistic regression model with only main effects
logit <- caret::train(as.factor(samesource) ~ ccf + rough_cor + D + sd_D + matches + 
                        mismatches + cms + non_cms + sum_peaks,
            data = hamby173and252_train, 
            method = "glm", 
            family = "binomial")

>>>>

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, eval = TRUE, 
                      cache = TRUE)
```

@@

```{r}
# Load libraries
library(lime)
library(tidyverse)
```

@@

```{r eval = FALSE}
# Fit a logistic regression model with only main effects
logit <- caret::train(as.factor(samesource) ~ ccf + rough_cor + D + sd_D + matches + 
                        mismatches + cms + non_cms + sum_peaks,
            data = hamby173and252_train, 
            method = "glm", 
            family = "binomial")
